---
title: "Bo Bichette In-Depth Analysis"
author: "Rayyan Haider"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 6
)
```

# Purpose

This document analyzes Bo Bichette's 2023 balls in play using a corrected and internally consistent coordinate system. The key correction is that the baseball diamond overlay is now drawn in the same transformed Statcast coordinate units as the batted-ball locations. This prevents second base, the mound, and the infield from being plotted unrealistically far into the outfield.

This version also separates **physical field-space coordinates** from **batter-perspective directional coordinates**. The field-space analyses use no handedness adjustment: KDE, 9-zone maps, Moran's I, LISA, Ripley's L, convex hull-style spatial summaries if added later, and entropy-style spatial summaries if added later should all use the unflipped field coordinates. Handedness adjustment is used only for spray-angle analysis, pull/center/opposite classification, mean spray angle, and directional statistics.

# Load Packages and Data

```{r libraries}
library(baseballr)
library(dplyr)
library(readr)
library(httr)
library(ggplot2)
library(tidyr)
library(knitr)
library(sf)
library(purrr)
```

```{r download-data}
# Bo Bichette MLBAM ID
bo_id <- 666182

# Direct Baseball Savant CSV URL for Bo Bichette 2023 batting Statcast data
url <- paste0(
  "https://baseballsavant.mlb.com/statcast_search/csv?",
  "all=true&hfPT=&hfAB=&hfBBT=&hfPR=&hfZ=&stadium=&hfBBL=&hfNewZones=&",
  "hfGT=R%7CPO%7CS%7C&hfC=&hfSea=2023%7C&hfSit=&hfOuts=&opponent=&",
  "pitcher_throws=&batter_stands=&hfSA=&player_type=batter&hfInfield=&",
  "team=&position=&hfOutfield=&hfRO=&home_road=&",
  "batters_lookup%5B%5D=", bo_id,
  "&game_date_gt=2023-03-30&game_date_lt=2023-10-01&hfFlag=&hfPull=&",
  "metric_1=&hfInn=&min_pitches=0&min_results=0&group_by=name&",
  "sort_col=pitches&player_event_sort=h_launch_speed&sort_order=desc&",
  "min_abs=0&type=details"
)

# Download to a temporary file
tmp <- tempfile(fileext = ".csv")
GET(url, write_disk(tmp, overwrite = TRUE))

# Read CSV directly
bo_raw <- read_csv(tmp, show_col_types = FALSE)

# Keep only balls in play with valid Statcast hit-coordinate values
bo_bip <- bo_raw %>%
  filter(type == "X", !is.na(hc_x), !is.na(hc_y)) %>%
  select(
    player_name,
    game_date,
    events,
    stand,
    batter,
    pitcher,
    hc_x,
    hc_y,
    launch_speed,
    launch_angle,
    hit_distance_sc,
    estimated_ba_using_speedangle,
    estimated_woba_using_speedangle
  )

# Save cleaned file for reproducibility
write_csv(bo_bip, "bo_bichette_2023_bip.csv")

nrow(bo_bip)
```

# Coordinate System and Field Geometry

The Statcast `hc_x` and `hc_y` values are not geographic coordinates and should not be treated as a GIS CRS. They are display-style hit-coordinate values. The analysis below centers home plate at approximately `(0, 0)` using the same no-handedness field-space transformation throughout the document:

\[
\text{field\_x} = hc_x - 125
\]

\[
\text{field\_y} = 200 - hc_y
\]

The baseball diamond is then scaled into these same coordinate units by dividing official baseball distances in feet by `2.5`. This makes second base, the pitcher's mound, and the other bases align with the transformed Statcast coordinate system rather than being plotted too deep into the outfield.

```{r coordinate-system-and-field-geometry}
# -----------------------------
# Coordinate transformation constants
# -----------------------------
HOME_PLATE_HC_X <- 125
HOME_PLATE_HC_Y <- 200
FEET_PER_STATCAST_UNIT <- 2.5

# Helper function: feet to transformed Statcast plotting units
ft_to_units <- function(feet) feet / FEET_PER_STATCAST_UNIT

# -----------------------------
# Create plotting dataset
# -----------------------------
bo_plot <- bo_bip %>%
  filter(!is.na(hc_x), !is.na(hc_y)) %>%
  mutate(
    # Physical field coordinates in transformed Statcast units.
    # IMPORTANT: these are NOT handedness-adjusted and should be used for
    # KDE, 9-zone maps, Moran's I, LISA, Ripley's L, convex hulls, and entropy.
    field_x = hc_x - HOME_PLATE_HC_X,
    field_y = HOME_PLATE_HC_Y - hc_y,

    # Backward-compatible aliases used by the plotting code below.
    # These remain physical field-space coordinates, not batter-perspective flips.
    spray_x = field_x,
    spray_y = field_y,

    # Batter-perspective x-coordinate used ONLY for directional statistics.
    # Right-handed batters keep the physical x-coordinate; left-handed batters are reflected.
    batter_x = if_else(stand == "L", -field_x, field_x),
    batter_y = field_y,

    result_class = case_when(
      events == "single" ~ "Single",
      events == "double" ~ "Double",
      events == "triple" ~ "Triple",
      events == "home_run" ~ "Home Run",
      events %in% c(
        "field_out", "force_out", "grounded_into_double_play",
        "double_play", "triple_play", "fielders_choice_out",
        "sac_fly", "sac_bunt", "other_out"
      ) ~ "Out",
      events == "field_error" ~ "Error",
      is.na(events) ~ "Other",
      TRUE ~ "Other"
    )
  )

# -----------------------------
# Field geometry in same units as field_x / field_y
# -----------------------------
base_90 <- ft_to_units(90)
first_base_center  <- c(base_90 / sqrt(2), base_90 / sqrt(2))
second_base_center <- c(0, base_90 * sqrt(2))
third_base_center  <- c(-base_90 / sqrt(2), base_90 / sqrt(2))

mound <- data.frame(
  x = 0,
  y = ft_to_units(60.5)
)

# helper function to make a small rotated square for each base
make_base <- function(cx, cy, s = ft_to_units(7), name = "base") {
  data.frame(
    x = c(cx, cx + s, cx, cx - s, cx),
    y = c(cy + s, cy, cy - s, cy, cy + s),
    base = name
  )
}

first_base <- make_base(first_base_center[1], first_base_center[2], name = "First")
second_base <- make_base(second_base_center[1], second_base_center[2], name = "Second")
third_base <- make_base(third_base_center[1], third_base_center[2], name = "Third")
bases_poly <- bind_rows(first_base, second_base, third_base)

home_plate <- data.frame(
  x = ft_to_units(c(-8.5 / 12, -8.5 / 12, 0, 8.5 / 12, 8.5 / 12, -8.5 / 12)),
  y = ft_to_units(c(0, 8.5 / 12, 17 / 12, 8.5 / 12, 0, 0))
)

# Foul lines and field arcs in same coordinate units
foul_line_distance <- ft_to_units(335)
center_field_distance <- ft_to_units(410)
infield_arc_radius <- ft_to_units(95)

foul_lines <- data.frame(
  x = c(0, -foul_line_distance / sqrt(2), 0, foul_line_distance / sqrt(2)),
  y = c(0,  foul_line_distance / sqrt(2), 0, foul_line_distance / sqrt(2)),
  grp = c("L", "L", "R", "R")
)

theta_in <- seq(pi / 4, 3 * pi / 4, length.out = 200)
infield_arc <- data.frame(
  x = infield_arc_radius * cos(theta_in),
  y = infield_arc_radius * sin(theta_in)
)

theta_out <- seq(pi / 4, 3 * pi / 4, length.out = 300)
outfield_arc <- data.frame(
  x = center_field_distance * cos(theta_out),
  y = center_field_distance * sin(theta_out)
)

# Plot limits based on both data and field geometry
plot_x_limits <- c(
  min(c(bo_plot$field_x, outfield_arc$x, foul_lines$x), na.rm = TRUE) - 5,
  max(c(bo_plot$field_x, outfield_arc$x, foul_lines$x), na.rm = TRUE) + 5
)
plot_y_limits <- c(
  0,
  max(c(bo_plot$field_y, outfield_arc$y, foul_lines$y), na.rm = TRUE) + 5
)

# A reusable field overlay layer function
add_field_overlay <- function() {
  list(
    geom_path(data = outfield_arc, aes(x = x, y = y), linewidth = 0.9, color = "black", inherit.aes = FALSE),
    geom_path(data = infield_arc, aes(x = x, y = y), linewidth = 0.9, color = "black", inherit.aes = FALSE),
    geom_line(data = filter(foul_lines, grp == "L"), aes(x = x, y = y), linewidth = 0.9, color = "black", inherit.aes = FALSE),
    geom_line(data = filter(foul_lines, grp == "R"), aes(x = x, y = y), linewidth = 0.9, color = "black", inherit.aes = FALSE),
    geom_polygon(data = bases_poly, aes(x = x, y = y, group = base), fill = "white", color = "black", linewidth = 0.9, inherit.aes = FALSE),
    geom_polygon(data = home_plate, aes(x = x, y = y), fill = "white", color = "black", linewidth = 0.9, inherit.aes = FALSE),
    geom_point(data = mound, aes(x = x, y = y), size = 2, color = "black", inherit.aes = FALSE)
  )
}
```

# Spray Chart

```{r spray-chart, fig.width=8, fig.height=6}
ggplot() +
  add_field_overlay() +
  geom_point(
    data = bo_plot,
    aes(x = field_x, y = field_y, color = result_class),
    alpha = 0.7,
    size = 2.2
  ) +
  coord_fixed(xlim = plot_x_limits, ylim = plot_y_limits, expand = FALSE) +
  labs(
    title = "Bo Bichette 2023 Spray Chart",
    subtitle = "Balls in play classified by result using unflipped physical field-space coordinates",
    x = NULL,
    y = NULL,
    color = "Result"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    legend.position = "right"
  )
```

# Kernel Density Estimation

```{r kde-map, fig.width=8, fig.height=6}
ggplot() +
  stat_density_2d(
    data = bo_plot,
    aes(x = field_x, y = field_y, fill = after_stat(level), alpha = after_stat(level)),
    geom = "polygon",
    contour = TRUE
  ) +
  add_field_overlay() +
  geom_point(
    data = bo_plot,
    aes(x = field_x, y = field_y),
    color = "black",
    alpha = 0.22,
    size = 1.1
  ) +
  coord_fixed(xlim = plot_x_limits, ylim = plot_y_limits, expand = FALSE) +
  scale_fill_viridis_c(option = "C", name = "Density") +
  scale_alpha(range = c(0.15, 0.7), guide = "none") +
  labs(
    title = "Bo Bichette 2023 Kernel Density Map",
    subtitle = "Density of balls in play with corrected baseball diamond overlay",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    legend.position = "right"
  )
```

# Radial 9-Zone Cluster Visualization and Chi-Square Test

The original rectangular 9-zone framework is replaced here with a baseball-specific radial classification system. Each batted ball is classified by **distance from home plate** and **spray direction** using the unflipped physical field-space coordinates. The three fixed distance bands are infield contact from 0--150 feet, shallow outfield contact from 150--250 feet, and deep outfield contact from 250--450 feet. Each band is then split into left field, center field, and right field using fixed angle ranges inside fair territory. This section is intentionally **not handedness-adjusted** because it describes where the balls landed on the field, not pull/center/opposite tendencies from the hitter's perspective.

```{r nine-zone-analysis, fig.width=8, fig.height=6}
# -----------------------------
# Radial fixed-distance 9-zone classification
# -----------------------------
# Depth bands are based on fixed baseball distances from home plate:
#   Infield:          0 to 150 feet
#   Shallow Outfield: 150 to 250 feet
#   Deep Outfield:    250 to 450 feet
#
# Directional bands are based on field angle:
#   Right Field: 45 to 75 degrees
#   Center:      75 to 105 degrees
#   Left Field:  105 to 135 degrees
#
# Caught foul balls are excluded from this specific 9-zone/entropy analysis
# because the 9-zone study window is defined over fair territory only.

zone_depth_breaks_ft <- c(0, 150, 250, 450)
zone_depth_breaks <- ft_to_units(zone_depth_breaks_ft)

zone_x_levels <- c("Left Field", "Center", "Right Field")
zone_y_levels <- c("Infield", "Shallow Outfield", "Deep Outfield")

zone_direction_specs <- tibble::tribble(
  ~x_zone,       ~angle_min_deg, ~angle_max_deg,
  "Right Field", 45,             75,
  "Center",      75,             105,
  "Left Field",  105,            135
) %>%
  mutate(x_zone = factor(x_zone, levels = zone_x_levels))

zone_depth_specs <- tibble::tribble(
  ~y_zone,             ~inner_ft, ~outer_ft,
  "Infield",           0,         150,
  "Shallow Outfield",  150,       250,
  "Deep Outfield",     250,       450
) %>%
  mutate(
    y_zone = factor(y_zone, levels = zone_y_levels),
    inner_r = ft_to_units(inner_ft),
    outer_r = ft_to_units(outer_ft)
  )

make_sector_polygon <- function(inner_r, outer_r, angle_min_deg, angle_max_deg, n = 120) {
  outer_theta <- seq(angle_min_deg, angle_max_deg, length.out = n) * pi / 180
  inner_theta <- seq(angle_max_deg, angle_min_deg, length.out = n) * pi / 180

  outer_coords <- cbind(
    outer_r * cos(outer_theta),
    outer_r * sin(outer_theta)
  )

  if (inner_r == 0) {
    coords <- rbind(
      c(0, 0),
      outer_coords,
      c(0, 0)
    )
  } else {
    inner_coords <- cbind(
      inner_r * cos(inner_theta),
      inner_r * sin(inner_theta)
    )
    coords <- rbind(
      outer_coords,
      inner_coords,
      outer_coords[1, ]
    )
  }

  st_polygon(list(coords))
}

zone_polygons_df <- expand_grid(
  y_zone = zone_depth_specs$y_zone,
  x_zone = zone_direction_specs$x_zone
) %>%
  left_join(zone_depth_specs, by = "y_zone") %>%
  left_join(zone_direction_specs, by = "x_zone") %>%
  mutate(
    zone_9 = paste(y_zone, x_zone, sep = "_"),
    geometry = pmap(
      list(inner_r, outer_r, angle_min_deg, angle_max_deg),
      ~ make_sector_polygon(..1, ..2, ..3, ..4)
    )
  )

zone_polygons <- st_sf(
  zone_polygons_df %>% select(-geometry),
  # These are custom baseball-field coordinates, not latitude/longitude or a projected map CRS.
  # Do not set crs = NA here; logical NA causes sf to throw an error.
  geometry = do.call(st_sfc, zone_polygons_df$geometry)
) %>%
  mutate(
    zone_area = as.numeric(st_area(geometry)),
    area_share = zone_area / sum(zone_area)
  )

bo_zones <- bo_plot %>%
  filter(!is.na(field_x), !is.na(field_y)) %>%
  mutate(
    radial_distance_units = sqrt(field_x^2 + field_y^2),
    radial_distance_ft = radial_distance_units * FEET_PER_STATCAST_UNIT,
    field_angle_deg = atan2(field_y, field_x) * 180 / pi,

    # Fair-territory radial window.
    in_fair_radial_window = field_angle_deg >= 45 &
      field_angle_deg <= 135 &
      radial_distance_ft >= zone_depth_breaks_ft[1] &
      radial_distance_ft <= zone_depth_breaks_ft[length(zone_depth_breaks_ft)],

    x_zone = case_when(
      field_angle_deg >= 105 & field_angle_deg <= 135 ~ "Left Field",
      field_angle_deg >= 75  & field_angle_deg < 105  ~ "Center",
      field_angle_deg >= 45  & field_angle_deg < 75   ~ "Right Field",
      TRUE ~ NA_character_
    ),
    y_zone = case_when(
      radial_distance_ft >= 0   & radial_distance_ft <= 150 ~ "Infield",
      radial_distance_ft > 150  & radial_distance_ft <= 250 ~ "Shallow Outfield",
      radial_distance_ft > 250  & radial_distance_ft <= 450 ~ "Deep Outfield",
      TRUE ~ NA_character_
    ),
    x_zone = factor(x_zone, levels = zone_x_levels),
    y_zone = factor(y_zone, levels = zone_y_levels),
    zone_9 = paste(y_zone, x_zone, sep = "_")
  ) %>%
  filter(in_fair_radial_window, !is.na(x_zone), !is.na(y_zone))

n_zone_included <- nrow(bo_zones)
n_zone_excluded <- bo_plot %>% filter(!is.na(field_x), !is.na(field_y)) %>% nrow() - n_zone_included

zone_counts_9 <- zone_polygons %>%
  st_drop_geometry() %>%
  select(y_zone, x_zone, zone_9, zone_area, area_share, inner_ft, outer_ft, angle_min_deg, angle_max_deg) %>%
  left_join(
    bo_zones %>% count(y_zone, x_zone, zone_9, name = "observed"),
    by = c("y_zone", "x_zone", "zone_9")
  ) %>%
  mutate(
    observed = ifelse(is.na(observed), 0, observed),
    observed_proportion = observed / sum(observed),
    expected = sum(observed) * area_share,
    density_ratio = observed_proportion / area_share,
    residual = observed - expected
  ) %>%
  arrange(
    factor(y_zone, levels = c("Deep Outfield", "Shallow Outfield", "Infield")),
    factor(x_zone, levels = c("Left Field", "Center", "Right Field"))
  )

kable(
  zone_counts_9 %>%
    select(y_zone, x_zone, observed, observed_proportion, area_share, expected, density_ratio),
  digits = 3
)

# Chi-square goodness-of-fit test
# Null hypothesis: balls in play are distributed across the 9 zones in proportion
# to the fair-territory area of each radial wedge.
chisq_result_9 <- chisq.test(
  x = zone_counts_9$observed,
  p = zone_counts_9$area_share,
  rescale.p = TRUE
)

chisq_result_9

expected_counts_9 <- zone_counts_9 %>%
  mutate(expected = round(expected, 2)) %>%
  select(zone_9, observed, expected, observed_proportion, area_share, density_ratio, residual)

kable(expected_counts_9, digits = 3)

# Entropy summaries
# Raw entropy describes how evenly Bo's batted balls are spread across the 9 zones.
# Area-adjusted entropy describes how evenly the distribution looks after accounting
# for how much fair-territory area each zone contains.
observed_p <- zone_counts_9$observed_proportion
area_p <- zone_counts_9$area_share
nonzero <- observed_p > 0 & area_p > 0

raw_entropy <- -sum(observed_p[observed_p > 0] * log(observed_p[observed_p > 0]))
normalized_raw_entropy <- raw_entropy / log(nrow(zone_counts_9))
kl_divergence_area <- sum(observed_p[nonzero] * log(observed_p[nonzero] / area_p[nonzero]))
area_adjusted_evenness <- exp(-kl_divergence_area)

entropy_summary_9 <- tibble(
  n_included = n_zone_included,
  n_excluded = n_zone_excluded,
  raw_entropy = raw_entropy,
  normalized_raw_entropy = normalized_raw_entropy,
  kl_divergence_from_area_baseline = kl_divergence_area,
  area_adjusted_evenness = area_adjusted_evenness
)

kable(entropy_summary_9, digits = 3)

zone_sf <- zone_polygons %>%
  left_join(
    zone_counts_9 %>% select(zone_9, observed, observed_proportion, expected, density_ratio),
    by = "zone_9"
  ) %>%
  mutate(
    observed = ifelse(is.na(observed), 0, observed),
    observed_proportion = ifelse(is.na(observed_proportion), 0, observed_proportion),
    expected = ifelse(is.na(expected), 0, expected),
    label = paste0(
      observed,
      "\nObs: ", round(100 * observed_proportion, 1), "%",
      "\nExp: ", round(100 * area_share, 1), "%"
    )
  )

label_points <- st_centroid(zone_sf) %>%
  mutate(
    label_x = st_coordinates(.)[, 1],
    label_y = st_coordinates(.)[, 2]
  ) %>%
  st_drop_geometry()

# Radial 9-zone cluster visualization
nine_zone_plot <- ggplot() +
  geom_sf(
    data = zone_sf,
    aes(fill = observed),
    color = "white",
    linewidth = 0.8,
    alpha = 0.75
  ) +
  add_field_overlay() +
  geom_point(
    data = bo_zones,
    aes(x = field_x, y = field_y),
    alpha = 0.25,
    size = 0.9,
    color = "black"
  ) +
  geom_text(
    data = label_points,
    aes(x = label_x, y = label_y, label = label),
    fontface = "bold",
    size = 3.1
  ) +
  coord_sf(xlim = plot_x_limits, ylim = plot_y_limits, expand = FALSE, datum = NA) +
  scale_fill_viridis_c(option = "C", name = "BIP Count") +
  labs(
    title = "Bo Bichette 2023 Radial 9-Zone Batted-Ball Map",
    subtitle = "Fixed-distance depth bands: infield 0-150 ft, shallow outfield 150-250 ft, deep outfield 250-450 ft; expected shares are area-weighted",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 9),
    legend.position = "right"
  )

nine_zone_plot
```

The chi-square test now evaluates whether Bichette's batted balls are distributed across the nine radial zones in proportion to the fair-territory area of each zone. This is more appropriate than testing against an even 1/9 baseline because the zones are not all the same size. A small p-value indicates that the observed distribution differs from the area-adjusted expected distribution.

# Local Moran's I with Area-Adjusted Rectangular Field Cells

The radial LISA grid is replaced here with a clipped rectangular field grid. This is easier to interpret for Local Moran's I because each spatial unit represents a small local area of the field, and neighboring cells are defined by ordinary grid adjacency. To avoid the original problem of cells extending into non-field space, the rectangular grid is clipped to the fair-territory study window. Local Moran's I is calculated on **area-adjusted standardized residuals**, not raw batted-ball counts. This means each cell's observed count is compared to the expected count based on that cell's share of the total included field area.

```{r local-morans-i-rectangular-area-adjusted, fig.width=8, fig.height=6}
library(sf)
library(spdep)

# -----------------------------
# Rectangular field grid for Local Moran's I
# -----------------------------
# This LISA version uses small rectangular/square cells clipped to the fair-territory
# field window. The expected count in each cell is proportional to the cell's clipped
# field area, so partial edge cells do not distort the analysis.
#
# Local Moran's I is run on the standardized area-adjusted residual:
#   (Observed count - Expected count) / sqrt(Expected count)
# where Expected count is based on each cell's share of total included field area.

# Study window: fair territory from the RF foul line to the LF foul line,
# using the same custom baseball-field coordinates as the rest of the project.
# These are not latitude/longitude coordinates and no geographic CRS is assigned.
lisa_field_radius <- ft_to_units(450)
lisa_theta <- seq(pi / 4, 3 * pi / 4, length.out = 400)

lisa_field_coords <- rbind(
  c(0, 0),
  cbind(lisa_field_radius * cos(lisa_theta), lisa_field_radius * sin(lisa_theta)),
  c(0, 0)
)

lisa_field_polygon <- st_sf(
  geometry = st_sfc(st_polygon(list(lisa_field_coords)))
)

# Create an evenly spaced rectangular grid, then clip it to fair territory.
# A 50-foot cell size gives a fine-grained map without making the cells too sparse.
lisa_cell_size <- ft_to_units(50)

lisa_grid_raw <- st_make_grid(
  lisa_field_polygon,
  cellsize = lisa_cell_size,
  square = TRUE
) %>%
  st_sf(geometry = .) %>%
  mutate(raw_cell_id = row_number())

lisa_grid_clipped <- suppressWarnings(
  st_intersection(lisa_grid_raw, lisa_field_polygon)
) %>%
  st_collection_extract("POLYGON") %>%
  mutate(
    lisa_cell_id = row_number(),
    cell_area = as.numeric(st_area(geometry)),
    full_cell_area = lisa_cell_size^2,
    area_fraction_of_full_cell = cell_area / full_cell_area
  ) %>%
  # Drop extremely tiny slivers that can create unstable expected counts.
  filter(area_fraction_of_full_cell >= 0.10) %>%
  mutate(lisa_cell_id = row_number())

# Points included in the LISA analysis are restricted to the same fair-territory
# window used by the clipped rectangular grid.
bo_lisa_points <- bo_plot %>%
  filter(!is.na(field_x), !is.na(field_y)) %>%
  mutate(
    radial_distance_ft = sqrt(field_x^2 + field_y^2) * FEET_PER_STATCAST_UNIT,
    field_angle_deg = atan2(field_y, field_x) * 180 / pi,
    in_lisa_window = field_angle_deg >= 45 &
      field_angle_deg <= 135 &
      radial_distance_ft >= 0 &
      radial_distance_ft <= 450
  ) %>%
  filter(in_lisa_window)

bo_lisa_points_sf <- bo_lisa_points %>%
  st_as_sf(coords = c("field_x", "field_y"), remove = FALSE)

# Count batted balls in each clipped rectangular cell.
lisa_point_counts <- st_join(bo_lisa_points_sf, lisa_grid_clipped, join = st_within) %>%
  st_drop_geometry() %>%
  filter(!is.na(lisa_cell_id)) %>%
  count(lisa_cell_id, name = "bip_count")

# Area-adjusted baseline and residuals.
lisa_zone_counts <- lisa_grid_clipped %>%
  left_join(lisa_point_counts, by = "lisa_cell_id") %>%
  mutate(
    bip_count = ifelse(is.na(bip_count), 0, bip_count),
    area_share = cell_area / sum(cell_area),
    expected_count = sum(bip_count) * area_share,
    observed_share = bip_count / sum(bip_count),
    observed_expected_ratio = ifelse(expected_count > 0, bip_count / expected_count, NA_real_),
    area_adjusted_residual = bip_count - expected_count,
    standardized_residual = ifelse(expected_count > 0, area_adjusted_residual / sqrt(expected_count), NA_real_),
    residual_centered = standardized_residual - mean(standardized_residual, na.rm = TRUE)
  )

# Spatial neighbors and row-standardized weights for the clipped rectangular cells.
# Queen contiguity is used so cells sharing either an edge or a corner count as neighbors.
lisa_neighbors <- poly2nb(lisa_zone_counts, queen = TRUE)
lisa_weights <- nb2listw(lisa_neighbors, style = "W", zero.policy = TRUE)

# Local Moran's I statistics on area-adjusted standardized residuals.
lisa_local <- localmoran(
  lisa_zone_counts$standardized_residual,
  lisa_weights,
  zero.policy = TRUE
)

lisa_zone_counts <- lisa_zone_counts %>%
  mutate(
    local_i = lisa_local[, "Ii"],
    local_i_z = lisa_local[, "Z.Ii"],
    local_i_p = lisa_local[, "Pr(z != E(Ii))"],
    spatial_lag_residual = lag.listw(lisa_weights, standardized_residual, zero.policy = TRUE),
    lag_centered = spatial_lag_residual - mean(standardized_residual, na.rm = TRUE),
    lisa_cluster = case_when(
      local_i_p <= 0.05 & residual_centered > 0 & lag_centered > 0 ~ "High-High",
      local_i_p <= 0.05 & residual_centered < 0 & lag_centered < 0 ~ "Low-Low",
      local_i_p <= 0.05 & residual_centered > 0 & lag_centered < 0 ~ "High-Low",
      local_i_p <= 0.05 & residual_centered < 0 & lag_centered > 0 ~ "Low-High",
      TRUE ~ "Not Significant"
    ),
    lisa_cluster = factor(
      lisa_cluster,
      levels = c("High-High", "Low-Low", "High-Low", "Low-High", "Not Significant")
    )
  )

# Table of the local statistics used in the LISA map.
lisa_zone_counts %>%
  st_drop_geometry() %>%
  select(
    lisa_cell_id,
    cell_area,
    area_fraction_of_full_cell,
    bip_count,
    expected_count,
    observed_expected_ratio,
    standardized_residual,
    spatial_lag_residual,
    local_i,
    local_i_z,
    local_i_p,
    lisa_cluster
  ) %>%
  arrange(lisa_cell_id) %>%
  kable(digits = 3)

# Plot limits based on the clipped rectangular LISA grid.
lisa_bbox <- st_bbox(lisa_zone_counts)
lisa_x_limits <- c(lisa_bbox["xmin"] - 5, lisa_bbox["xmax"] + 5)
lisa_y_limits <- c(0, lisa_bbox["ymax"] + 5)

# Local Moran's I cluster map using clipped rectangular cells and area-adjusted residuals.
ggplot() +
  geom_sf(
    data = lisa_zone_counts,
    aes(fill = lisa_cluster),
    color = "white",
    linewidth = 0.35,
    alpha = 0.8
  ) +
  geom_point(
    data = bo_lisa_points,
    aes(x = field_x, y = field_y),
    alpha = 0.3,
    size = 0.7
  ) +
  add_field_overlay() +
  coord_sf(xlim = lisa_x_limits, ylim = lisa_y_limits, expand = FALSE, datum = NA) +
  scale_fill_manual(
    values = c(
      "High-High" = "#F8766D",
      "Low-Low" = "#00BA38",
      "High-Low" = "#C77CFF",
      "Low-High" = "#619CFF",
      "Not Significant" = "#4D89F9"
    ),
    drop = FALSE,
    name = "LISA Cluster"
  ) +
  labs(
    title = "Bo Bichette 2023 Local Moran's I Cluster Map",
    subtitle = "Local spatial autocorrelation of area-adjusted BIP residuals using clipped rectangular field cells",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 9),
    legend.position = "right"
  )
```

This LISA analysis uses small rectangular cells clipped to fair territory and evaluates clustering in **area-adjusted residuals**, not raw counts. High-High cells identify neighboring field areas where Bichette recorded more batted balls than expected given cell area, while Low-Low cells identify neighboring areas with fewer batted balls than expected. High-Low and Low-High cells identify local spatial outliers. The global Moran's I grid map remains excluded because the local cluster map is more useful for identifying where Bichette's batted-ball concentrations occur on the field.


# Ripley's L Function with Monte Carlo Envelope

```{r ripleys-l, fig.width=8, fig.height=6}
library(spatstat.geom)
library(spatstat.explore)

bo_xy <- bo_plot %>%
  filter(!is.na(field_x), !is.na(field_y)) %>%
  select(field_x, field_y)

# Fair-territory window using the same coordinate units as the KDE and LISA analysis.
# These are custom baseball-field coordinates, not a geographic CRS.
field_radius <- max(
  center_field_distance,
  sqrt(bo_plot$field_x^2 + bo_plot$field_y^2),
  na.rm = TRUE
) + 5

theta_field <- seq(pi / 4, 3 * pi / 4, length.out = 300)

field_window_coords <- data.frame(
  x = c(0, field_radius * cos(theta_field), 0),
  y = c(0, field_radius * sin(theta_field), 0)
)

field_window <- owin(
  poly = list(
    x = field_window_coords$x,
    y = field_window_coords$y
  )
)

inside_field <- inside.owin(
  x = bo_xy$field_x,
  y = bo_xy$field_y,
  w = field_window
)

bo_xy_in <- bo_xy[inside_field, ]

bo_ppp <- ppp(
  x = bo_xy_in$field_x,
  y = bo_xy_in$field_y,
  window = field_window
)

summary(bo_ppp)

set.seed(123)

bo_l_envelope <- envelope(
  bo_ppp,
  fun = Lest,
  nsim = 99,
  correction = "border",
  global = FALSE
)

bo_l_envelope_df <- data.frame(
  distance = bo_l_envelope$r,
  observed_L = bo_l_envelope$obs,
  theoretical_L = bo_l_envelope$theo,
  lower = bo_l_envelope$lo,
  upper = bo_l_envelope$hi
)

ggplot(bo_l_envelope_df, aes(x = distance)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.25) +
  geom_line(aes(y = observed_L), linewidth = 1) +
  geom_line(aes(y = theoretical_L), linetype = "dashed", linewidth = 1) +
  labs(
    title = "Bo Bichette 2023 Ripley's L Function with Monte Carlo Envelope",
    subtitle = "Observed clustering across distances compared to simulated complete spatial randomness",
    x = "Distance in transformed Statcast coordinate units",
    y = "L(r)"
  ) +
  theme_minimal()
```

# Directional Spray Angle Analysis

This section is the only part of the file that uses batter-perspective handedness adjustment. The angle is calculated from `batter_x` and `batter_y`, where left-handed hitters would be horizontally reflected and right-handed hitters, including Bichette, remain unchanged. Negative angles indicate pull-side contact, positive angles indicate opposite-field contact, and values near zero indicate center-field contact.

```{r directional-statistics}
bo_angles <- bo_plot %>%
  filter(!is.na(field_x), !is.na(field_y)) %>%
  mutate(
    raw_spray_angle_rad = atan2(batter_x, batter_y),
    spray_angle_deg = raw_spray_angle_rad * 180 / pi,
    spray_direction_model = case_when(
      spray_angle_deg < -15 ~ "Pull",
      spray_angle_deg >= -15 & spray_angle_deg <= 15 ~ "Center",
      spray_angle_deg > 15 ~ "Opposite",
      TRUE ~ NA_character_
    )
  )

bo_angles %>%
  summarize(
    mean_spray_angle = mean(spray_angle_deg, na.rm = TRUE),
    median_spray_angle = median(spray_angle_deg, na.rm = TRUE),
    sd_spray_angle = sd(spray_angle_deg, na.rm = TRUE),
    n = n()
  ) %>%
  kable(digits = 3)

bo_direction_model_summary <- bo_angles %>%
  count(spray_direction_model, name = "n") %>%
  mutate(percent = round(100 * n / sum(n), 1))

kable(bo_direction_model_summary, digits = 1)
```

```{r spray-angle-density, fig.width=8, fig.height=5}
ggplot(bo_angles, aes(x = spray_angle_deg)) +
  geom_density(linewidth = 1.2) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1) +
  labs(
    title = "Bo Bichette 2023 Spray Angle Density",
    subtitle = "Negative values indicate pull-side contact; positive values indicate opposite-field contact",
    x = "Spray Angle (degrees)",
    y = "Density"
  ) +
  theme_minimal()
```

```{r spray-direction-bar-chart, fig.width=7, fig.height=5}
ggplot(
  bo_direction_model_summary,
  aes(x = spray_direction_model, y = percent)
) +
  geom_col() +
  labs(
    title = "Bo Bichette 2023 Modeled Spray Direction",
    subtitle = "Classification based on spray-angle cutoffs",
    x = "Spray Direction",
    y = "Percent of Balls in Play"
  ) +
  theme_minimal()
```

```{r one-sample-spray-angle-test}
# Null hypothesis: mean spray angle = 0 degrees
# Interpretation: a mean of 0 represents a perfectly centered average spray direction.
bo_angle_ttest <- t.test(
  bo_angles$spray_angle_deg,
  mu = 0
)

bo_angle_ttest
```

# Notes for the 15-Player Extension

For the 15-player comparative file, the same physical coordinate pipeline should be used first. Then a separate handedness-adjusted batter-perspective coordinate should be created only for directional comparisons. Right-handed hitters keep the same field-space x-coordinate, while left-handed hitters should have the field-space x-coordinate multiplied by -1. The y-coordinate remains unchanged.

Physical field analyses such as KDE, 9-zone maps, Moran's I, LISA, Ripley's L, convex hulls, and entropy should use the physical `field_x` and `field_y` coordinates with no handedness adjustment. Batter-perspective spray angle summaries, pull/center/opposite classification, mean spray angle, and directional statistics should use `batter_x` and `batter_y`.
